---
title: "MS Keyrun Data Needs"
author: "Sarah Gaichas"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Decisions from meeting June 3 2020

Notes were taken on the [KEYRUN Kanban board](https://www.st.nmfs.noaa.gov/jira/secure/RapidBoard.jspa?rapidView=384&projectKey=KEYRUN) under the epic [KEYRUN-4](https://www.st.nmfs.noaa.gov/jira/browse/KEYRUN-4) (a password is required to access these sites, but current information is copied below).

## Objective: Develop a common input dataset for use in all keyrun models

The purpose of the June 3 meeting was to describe the decisions and components required to develop the common input dataset for the project.

### Epic description

(This text copied from KEYRUN-4 Description and lightly edited)

Below is a list of potential stories under this Epic with links to other Epics. Numbers indicate order tasks need to be done (priority) established at the June 3 2020 meeting. 

TODO: Add each story. Add done condition for each story. 

1. Document decisions on all stories below (maybe rmd) and document as we go for all datasets, Jira not searchable so establish decision document and documentation elsewhere (pull comment from done column? plugins may be available by license) 

1. Establish temporal and spatial dimensions for data (years, seasons, areas).

  define Georges Bank for survey and fisheries which will not match, could use habitat model

2. Establish species lists (Rpath KEYRUN-2 larger list than multispecies models KEYRUN-1 and KEYRUN-3)

2. Define environmental data needed for each model

2. Establish dataset standards (See e.g. condition work)

3. Biomass index data from surveys for all models:

4. Establish common definitions of fleets: applies to all catch data, discard data, and composition data.

5. Catch (landings) data for all models:

   Rpath in KEYRUN-2

   Hydra and MSCAA in KEYRUN-1

   All other tested models in KEYRUN-3

5. Discard data for all models, needs methods discussion/decision (check: Andy)

6. Length composition data from surveys and fisheries for Hydra in KEYRUN-1 (check: Andy)

6. Age composition data from surveys and fisheries for MSCAA in KEYRUN-1

6. Define which diet and or consumption data from surveys and fisheries for all models

6. Define which life history data from surveys and/or fisheries for all models (e.g. maturity, weight at age, maximum age for mortality estimation, etc)

 

### Additional meeting notes

Progress on catch data:

Notes on mscatch and comlands. comlands in a package now. dataset from comland is input to mscatch which combines with observer data length samples, for each quarter/yr for species, gear, can assign length samples to each of those landings bins. done and tested. Discards not yet accounted for, difficult and needs decisions. mostly documented, decision process for filling blanks documented in log file. Canadian landings for GB may be needed.

there is a comdisc as well, computes discard to kept ratio. add to subtasks under discard story

Identify done criteria, meet them, then run by key stock assessment folks. 

### Progress

Stories have been added to the board as of `r format(Sys.time(), '%d %B, %Y')`:

```{r, echo=FALSE}
htmltools::includeHTML("NMFS_JIRA_2020-06-15T_0_11_28-0400.html")
```

 
