---
title: "Biomass and Abundance by length"
output:
  html_document:
    code_fold: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, functions, echo = F}
catHeader <- function(text = "", level = 3) {
  cat(paste0("\n\n", 
               paste(rep("#", level), collapse = ""), 
               " ", text, "\n"))
}
```

The length based multispecies model (Hydra) requires annual indices of biomass and abundance by species size class. The NEFSC spring and fall [bottom trawl surveys](https://noaa-edab.github.io/tech-doc/) collect the information required to obtain estimates of these indices

Each tow/haul during the bottom trawl survey collects the total weight of each fish species caught. A sample of the haul is then taken and each fish in this sample is identified, lengthed (to the nearest cm) and weighed. 

The R package [`survdat`](https://noaa-edab.github.io/survdat/) is used to read in the survey data and provide estimates of swept area biomass and abundance on the Georges Bank footprint. The survey strata that define [Georges Bank](https://noaa-edab.github.io/ms-keyrun/articles/GBSurveySet.html) are 1090, 1130:1210, 1230, 1250, 3460, 3480, 3490, 3520:3550.


``` {r, echo = T}
source(here::here("data-raw/R/create_real_survey_lennumcomp.r"))
suppressMessages(out <- create_real_survey_lennumcomp(convertKGtoMT=T))


```

## Spring Biomass: length-frequency plots {.tabset}

```{r biomassS, results = "asis", echo = FALSE}
for(i in seq_along(out$plistBS)){
    tmp <- out$plistBS[[i]]
    # As you want to use tabset level here has to be lower than 
    # parent level (ie, parent is 2, so here you have to use 3)
    catHeader(tmp[[2]], 3)
    lapply(tmp[1], print)
}
```


## Fall Biomass: length-frequency plots {.tabset}

```{r biomassF, results = "asis", echo = FALSE}
for(i in seq_along(out$plistBF)){
    tmp <- out$plistBF[[i]]
    # As you want to use tabset level here has to be lower than
    # parent level (ie, parent is 2, so here you have to use 3)
    catHeader(tmp[[2]], 3)
    lapply(tmp[1], print)
}
```

## Spring Abundance: length-frequency plots {.tabset}

```{r abundS, results = "asis", echo = FALSE}
for(i in seq_along(out$plistAS)){
    tmp <- out$plistAS[[i]]
    # As you want to use tabset level here has to be lower than 
    # parent level (ie, parent is 2, so here you have to use 3)
    catHeader(tmp[[2]], 3)
    lapply(tmp[1], print)
}
```


## Fall Abundance: length-frequency plots {.tabset}

```{r abundF, results = "asis", echo = FALSE}
for(i in seq_along(out$plistAF)){
    tmp <- out$plistAF[[i]]
    # As you want to use tabset level here has to be lower than
    # parent level (ie, parent is 2, so here you have to use 3)
    catHeader(tmp[[2]], 3)
    lapply(tmp[1], print)
}
```


