---
title: "Landings and Length data sources"
output: html_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

All length and landings data were pulled from the NEFSC commerical fisheries database. Specifically the  stockeff.mv_cf_len and mv_cf_landings (since `comlandr` doesn't currently pull data in format required for mscatch. Ultimately the data will be pulled using the `comlandr` package). The data were saved as rda files in the `data` folder as:

* sampleData_itis_GB - for landings data
* sampleLengths_itis_GB - for length data

```{r echo = T, eval = F}
channel <- dbutils::connect_to_database("server","user")
test_data_pull_hydra_GB(channel)

```

The footprint used to pull the landings and length data were the statistical areas 521, 522, 523, 524, 525, 526, 538, 551, 552, 561, 562, 537 which comprise Georges Bank.

Herring and winter Skate were handled differently due to data availability.

Note that discards are not included in the sampleData (with the exception of Herring). This will be included at a later date.

## Herring

Herring data is poorly represented in NEFSC fisheries database so all landings and length data were pulled from the state of Maine. In the state of Maine database catch is reported in metric tons. This was converted to lbs to align with NEFSC commercial fisheries data base. Fish lengths are reported in mm. These were converted to cm to align with the NEFSC survey database

```{r, echo = T, eval = F}
get_herring_data()
format_herring_data()
```

## Winter Skate

No skate species are recorded in the NEFSC commercial fisheries database. All species are recorded as an unclassified skate. The package `comlandr` calculates the proportions of skate species in the surveys as applied them to the commercial landings to obtain estimates of winter skate. The lengths for winter skate were pulled from the bottom trawl survey. 

```{sql wslengths, echo = T, eval = F}
get_skate_data()
format_skate_data()
```



