---
title: "Georges Bank Survey Biomass and Catch Data"
output:
  html_document:
    code_fold: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)

library(dplyr)
library(ggplot2)
library(ggthemes)
library(ggiraph)
library(here)
```


All models use a fishery independent index of biomass and a catch time series. This document visualizes these basic assessment inputs.

## Fishery independent survey indices

The NEFSC spring and fall [bottom trawl surveys](https://noaa-edab.github.io/tech-doc/) collect the information required to obtain estimates of survey biomass and abundance. R package [`survdat`](https://noaa-edab.github.io/survdat/) was used to pull and compile the survey data time series. NEFSC bottom-trawl sampling encompasses about 293,000 square km of continental shelf from Cape Hatteras, NC, to Nova Scotia, Canada in depths from 8-400 m. 

Multiple survey datasets are provided in the `mskeyrun` package; all were created using `create_survey_index()` in `ms-keyrun/data-raw`. Datasets for the 10 Georges Bank focal species are `mskeyrun::surveyIndex` (all survey years 1968-2019 converted to survey vessel R/V Albatross units for years surveyed by R/V Henry Bigelow 2009-2019), and unconverted separate indices for Albatross years (`mskeyrun::surveyIndexA4`) and Bigelow years (`mskeyrun::surveyIndexHB`). Datasets for food web modeling include all surveyed species for combined and separate surveys. These datasets have the same names with the suffix -All appended (e.g. `mskeyrun::surveyIndexAll`)

Here we visualize survey indices from the `mskeyrun::surveyIndex` dataset.

```{r biomass}

modyears <- 1968:2019  # 1968 agreed for project, testing with shorter dataset
  
focalspp <- mskeyrun::focalSpecies %>%
      dplyr::filter(modelName != "Pollock") %>% # not using in these models
      dplyr::mutate(Name = modelName) 
    
    # single index, use if fitting to two not working
survindex.all <- mskeyrun::surveyIndex %>%
      dplyr::filter(YEAR %in% modyears) %>%
      dplyr::left_join(focalspp %>% dplyr::select(-NESPP3) %>% dplyr::distinct()) %>%
      dplyr::filter(!is.na(modelName)) %>%
      dplyr::mutate(ModSim = "Actual",
                    year = YEAR,
                    Code = SPECIES_ITIS,
                    Name = modelName,
                    survey = paste0("Combined ", SEASON)) %>%
      dplyr::select(ModSim, year, Code, Name, survey, variable, value, units)
    
towarea <- 0.0384
    
GBsurvstrata  <- c(1090, 1130:1210, 1230, 1250, 3460, 3480, 3490, 3520:3550)
    
GBarea <- FishStatsUtils::northwest_atlantic_grid %>%
      dplyr::filter(stratum_number %in% GBsurvstrata) %>%
      dplyr::summarise(area = sum(Area_in_survey_km2)) %>%
      as.double()

survbio <- survindex.all %>%
      dplyr::filter(variable %in% c("strat.biomass", "biomass.var")) %>%
      dplyr::select(-units) %>%
      tidyr::pivot_wider(names_from = "variable", values_from = "value") %>%
      dplyr::mutate(cv = sqrt(biomass.var)/strat.biomass) %>%
      dplyr::select(-biomass.var) %>%
      # expand: area of tow =  0.0384 and GB area = above and kg to tons
      dplyr::mutate(biomass = strat.biomass * (GBarea/towarea) /1000) %>% 
      tidyr::pivot_longer(c(biomass, cv), names_to = "variable", values_to = "value") %>%
      dplyr::mutate(units = ifelse(variable=="biomass", "tons", "unitless"))

# plot biomass time series facet wrapped by species
plotB <- function(dat, truedat=NULL){
  
    svbio <- dat %>% filter(variable=="biomass")
    svcv <- dat %>% filter(variable=="cv")
  
    ggplot() +
    geom_line(data=svbio, aes(x=year,y=value, color="Survey Biomass"), 
              alpha = 10/10) +
    {if(!is.null(truedat)) geom_line(data=truedat, aes(x=time/365,y=atoutput, color="True B"), alpha = 3/10)} + 
    theme_tufte() +
    theme(legend.position = "top") +
    xlab("model year") +
    ylab("tons") +
    labs(colour=dat$ModSim) +
    facet_wrap(~Name, scales="free") 
  
}
```

```{r, results="asis"}
 for(s in unique(survbio$survey)){
   cat("  \n##### ",  s,"  \n")
   print(plotB(survbio %>%
                 filter((survey %in% s))))
   cat("  \n")
 }

```


## Fishery dependent catch 

Stuff about where this comes from

Separate discards and landings

Temporary total catch here

```{r catch, results = "asis", echo = FALSE}

    fleetdef <- focalspp %>%
      dplyr::select(-NESPP3) %>% 
      dplyr::distinct() %>%
      dplyr::mutate(species = Name) %>%
      dplyr::arrange(species) %>%
      dplyr::mutate(pelagic = dplyr::case_when(species %in% c("Atlantic_herring",
                                                              "Atlantic_mackerel") ~ 2,
                                               TRUE ~ 0),
                    demersal = dplyr::case_when(!species %in% c("Atlantic_herring",
                                                                "Atlantic_mackerel") ~ 1,
                                                TRUE ~ 0),
                    qind = pelagic + demersal,
                    pelagic = pelagic/2) %>%
      dplyr::select(species, demersal, pelagic, qind)
    
    # need to have landings + discards = catch
    fishindex <- mskeyrun::catchIndex %>%
      dplyr::filter(YEAR %in% modyears) %>%
      dplyr::left_join(focalspp %>% dplyr::mutate(NESPP3 = as.integer(NESPP3))) %>%
      dplyr::filter(!is.na(Name)) %>%
      dplyr::select(-units) %>%
      tidyr::pivot_wider(names_from = "variable", values_from = "value") %>%
      dplyr::group_by(YEAR, Name) %>%
      dplyr::mutate(catch = sum(`commercial landings`,`commercial discards`, na.rm = TRUE),
                    cv = 0.05) %>%
      dplyr::ungroup() %>%
      tidyr::pivot_longer(c(`commercial landings`, `commercial discards`, catch, cv), names_to = "variable", values_to = "value") %>%
      dplyr::mutate(units = ifelse(variable=="cv", "unitless", "metric tons")) %>%
      dplyr::left_join(fleetdef, by=c("Name" = "species")) %>%
      dplyr::mutate(ModSim = "Actual",
                    year = YEAR,
                    Code = SPECIES_ITIS,
                    Name = modelName,
                    fishery = qind) %>%
      dplyr::select(ModSim, year, Code, Name, fishery, variable, value, units)

# make a catch series function that can be split by fleet? this doesnt
# also note different time (days) from model timestep in all other output
plotC <- function(dat, truedat=NULL){
  
    ctbio <- dat %>% filter(variable=="catch")
    ctcv <- dat %>% filter(variable=="cv")
  
    ggplot() +
    geom_line(data=ctbio, aes(x=year,y=value, color="Catch biomass"), 
              alpha = 10/10) +
    {if(!is.null(truedat)) geom_line(data=truedat, aes(x=time/365,y=atoutput, color="True Catch"), alpha = 3/10)} + 
    theme_tufte() +
    theme(legend.position = "top") +
    xlab("model year") +
    ylab("tons") +
    labs(colour=dat$ModSim) +
    facet_wrap(~Name, scales="free") 
  
}

plotC(fishindex)

```

